---
title: "Analyze HMDA Mortgage Data with Snowflake Cortex AI and Positron Assistant"
format: html
editor_options:
  chunk_output_type: console
execute:
  echo: false
  warning: false
  error: false
---

The Snowflake guide this document accompanies:

- Uses the Home Mortgage Disclosure Act (HMDA) dataset from Snowflake's free public data
- Leverages Cortex AI with Databot and Positron Assistant for exploratory data analysis in Posit Workbench
- Builds an interactive Shiny dashboard with traditional UI controls
- Deploys to Posit Connect for easy and secure sharing and collaboration

## Install packages

```{r}
install.packages("renv")
renv::restore()
```

## Connect to data

```{r}
library(DBI)
library(odbc)
library(dplyr)
library(dbplyr)
library(connectcreds)

get_connection <- function() {
  warehouse <- "DEFAULT_WH"
  database <- "SNOWFLAKE_PUBLIC_DATA_FREE"
  schema <- "PUBLIC_DATA_FREE"
  account <- Sys.getenv("SNOWFLAKE_ACCOUNT")

  con <- DBI::dbConnect(
    odbc::snowflake(),
    account = account,
    warehouse = warehouse,
    database = database,
    schema = schema
  )

  return(con)
}
```

## Exploratory data analysis with Databot

Let's step out of our Quarto document and open Databot to explore our mortgage data. Come back when you're ready to build the dashboard.

## Build Interactive Dashboard

### Build a Shiny app

```{r}
# Create app.R file with Shiny application code
app_code <- '
library(shiny)
library(DBI)
library(odbc)
library(dplyr)
library(dbplyr)
library(connectcreds)
library(ggplot2)

get_connection <- function() {
  warehouse <- "DEFAULT_WH"
  database <- "SNOWFLAKE_PUBLIC_DATA_FREE"
  schema <- "PUBLIC_DATA_FREE"
  account <- Sys.getenv("SNOWFLAKE_ACCOUNT")

  con <- DBI::dbConnect(
    odbc::snowflake(),
    account = account,
    warehouse = warehouse,
    database = database,
    schema = schema
  )

  return(con)
}

# Define UI
ui <- fluidPage(
  titlePanel("HMDA Mortgage Data Explorer"),
  sidebarLayout(
    sidebarPanel(
      selectInput("loan_type", "Select Loan Type:",
                  choices = c("All", "Conventional", "FHA", "VA", "FSA/RHS")),
      selectInput("state", "Select State:",
                  choices = c("All", "CA", "TX", "FL", "NY", "PA")),
      sliderInput("loan_amount", "Loan Amount Range:",
                  min = 0, max = 1000000, value = c(0, 500000),
                  step = 50000)
    ),
    mainPanel(
      h4("Loan Distribution:"),
      plotOutput("loan_plot"),
      h4("Summary Statistics:"),
      verbatimTextOutput("summary_stats")
    )
  )
)

# Define server
server <- function(input, output, session) {
  # Initialize connection
  con <- get_connection()
  mortgage_data <- tbl(con, "HOME_MORTGAGE_DISCLOSURE_ATTRIBUTES")

  # Reactive filtered data
  filtered_data <- reactive({
    data <- mortgage_data

    if (input$loan_type != "All") {
      data <- data %>% filter(LOAN_TYPE == input$loan_type)
    }

    if (input$state != "All") {
      data <- data %>% filter(STATE == input$state)
    }

    data %>%
      filter(LOAN_AMOUNT >= input$loan_amount[1],
             LOAN_AMOUNT <= input$loan_amount[2])
  })

  output$loan_plot <- renderPlot({
    data <- filtered_data() %>% collect()

    ggplot(data, aes(x = LOAN_AMOUNT)) +
      geom_histogram(bins = 30, fill = "steelblue", color = "white") +
      theme_minimal() +
      labs(title = "Distribution of Loan Amounts",
           x = "Loan Amount ($)",
           y = "Count")
  })

  output$summary_stats <- renderPrint({
    data <- filtered_data() %>% collect()
    summary(data$LOAN_AMOUNT)
  })
}

# Run the application
shinyApp(ui = ui, server = server)
'

# Write the app code to app.R file
writeLines(app_code, "app.R")

message("Shiny app created successfully!")
message("\nRun the app with: shiny::runApp(\"app.R\")")
```
