---
title: "Analyze HMDA Mortgage Data with Snowflake Cortex AI and Positron Assistant"
format: html
editor_options:
  chunk_output_type: console
execute:
  echo: false
  warning: false
  error: false
---

The Snowflake guide this document accompanies:

- Uses the Home Mortgage Disclosure Act (HMDA) dataset from Snowflake's free public data
- Leverages Cortex AI, Databot, Positron Assistant, {ellmer}, and {querychat} for LLM interactions in Posit Workbench
- Deploys to Posit Connect for easy and secure sharing and collaboration

## Install packages

```{r}
install.packages("renv")
renv::restore()
```

## Connect to data

```{r}
library(DBI)
library(odbc)
library(dplyr)
library(dbplyr)
library(connectcreds)

get_connection <- function() {
  warehouse <- "MORTGAGE_DATA_WH"
  database <- "SNOWFLAKE_PUBLIC_DATA_FREE"
  schema <- "PUBLIC_DATA_FREE"
  account <- Sys.getenv("SNOWFLAKE_ACCOUNT")

  con <- DBI::dbConnect(
    odbc::snowflake(),
    account = account,
    warehouse = warehouse,
    database = database,
    schema = schema
  )

  return(con)
}
```

## Exploratory data analysis with Databot

Let's step out of our Quarto document and open Databot to explore our mortgage data. Come back when you're ready to build the dashboard.

## Build LLM Dashboard

### Test {ellmer} connection

```{r}
library(ellmer)

# Initialize chat with Snowflake Cortex AI

chat <- chat(
  name = "snowflake/claude-sonnet-4-5",
  system_prompt = "You are a mortgage lending and housing finance data analysis expert",
  account = Sys.getenv("SNOWFLAKE_ACCOUNT")
)
# Test the connection
response <- chat$chat("What patterns do you see in home mortgage lending data?")
cat(response)
```

### Optional: Explore with {querychat}


**Note:** `querychat_app()` launches an interactive application that will block further code execution. Close the app and stop the function when you're done exploring to continue with the next steps.

```{r}
library(querychat)

querychat_app(
  data = mortgage_data,
  client = chat  # Use the Snowflake Cortex AI chat configured with ellmer
)
```


### Build a Shiny app

```{r}
# Create app.R file with Shiny application code
app_code <- '
library(shiny)
library(ellmer)
library(DBI)
library(odbc)
library(dplyr)
library(dbplyr)
library(connectcreds)

get_connection <- function() {
  warehouse <- "MORTGAGE_DATA_WH"
  database <- "SNOWFLAKE_PUBLIC_DATA_FREE"
  schema <- "PUBLIC_DATA_FREE"
  account <- Sys.getenv("SNOWFLAKE_ACCOUNT")

  con <- DBI::dbConnect(
    odbc::snowflake(),
    account = account,
    warehouse = warehouse,
    database = database,
    schema = schema
  )

  return(con)
}

# Define UI
ui <- fluidPage(
  titlePanel("HMDA Data Explorer with Snowflake Cortex AI"),
  sidebarLayout(
    sidebarPanel(
      textAreaInput("user_question", "Ask a question about mortgage data:",
                    placeholder = "What are the most common loan types?"),
      actionButton("ask", "Ask Cortex AI")
    ),
    mainPanel(
      h4("AI Response:"),
      verbatimTextOutput("response"),
      h4("Data Results:"),
      tableOutput("results")
    )
  )
)

# Define server
server <- function(input, output, session) {
  # Initialize connection and chat
  con <- get_connection()
  mortgage_data <- tbl(con, "HOME_MORTGAGE_DISCLOSURE_ATTRIBUTES")

  chat <- chat(
    name = "snowflake/claude-sonnet-4-5",
    system_prompt = "You are a mortgage lending and housing finance data analysis expert",
    account = Sys.getenv("SNOWFLAKE_ACCOUNT")
  )

  observeEvent(input$ask, {
    req(input$user_question)

    output$response <- renderText({
      chat$chat(input$user_question)
    })

    # Example: Could execute data queries based on the question
    # This would require parsing the question or using AI to generate queries
  })
}

# Run the application
shinyApp(ui = ui, server = server)
'

# Write the app code to app.R file
writeLines(app_code, "app.R")

message("Shiny app created successfully!")
message("- Cortex AI Model: snowflake/claude-sonnet-4-5")
message("\nRun the app with: shiny::runApp('app.R')")
```
